Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
text2

We describe and evaluate two approaches for synthesising digital controllers 
for state-space physical plants. Both approaches consider a counterexample 
guided inductive synthesis (CEGIS) loop. In the first approach, 
we devise a digital controller that stabilizes the system whilst remaining safe given a 
pre-selected time horizon; then, we verify its 
potential safety by unfolding the dynamics of the system and checking for 
a completeness threshold. In the second approach, we substitute the second step by employing abstract acceleration
to unfold the dynamics forward, instead of symbolically unwinding the 
transition relation~\cite{cattaruzza2015unbounded}. 
\todo{We need to mention the CEGIS architecture here. We need to emphasise the derivation of errors for soundness. Can we claim completeness already?}
We provide experimental results 
showing that both approaches are able to efficiently synthesise stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is \textcolor{red}{$XXX$}s, i.e., most controllers can be synthesized in less
than \textcolor{red}{$XXX$} seconds.

\textit{Contributions}. 
\begin{enumerate}
\item In this paper we extend the problem of digital control synthesis from a transfer function representation into a state space representation.
This entails the verification of safety properties in addition to the already existing stability conditions which is a much harder problem.
\item To achieve this, we develop two approaches and compare them, the first relying on a non-deterministic simulation of the system and a completeness proof, whilst the second leverages on the concepts of abstraction and acceleration to increase speed whilst retaining soundness.
\item We develop a solution generalization algorithm for the first approach and an abstraction refinement for the second, both of which improve the performance of the CEGIS loop. 
\item We develop a model for different sources quantization and their effect on the state space progression, for which we find safe bounds that ensure the safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback architectures can be employed to affect the properties and behaviours of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback, 
as implemented on modern FPGAs or ICs. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
The gain matrix $K$ can be set so that the closed-loop digitalized dynamics are shaped as desired, 
for instance according to a specific stability goal or around a specific dynamical behaviour \cite{astrom1997computer}. 
As argued later in this work, we will target more complex objectives, such as quantitative safety requirements that are not typical in the digital control literature.

@@

A discrete-time system as \eqref{eq:closedloopss} is asymptotically stable 
\textit{iff} all the roots of the characteristic polynomial 
(i.e., the eigenvalues of the closed-loop matrix $A_d - B_d K$) 
are inside the unity circle in the complex plane, i.e., \textit{iff} their absolute 
values are less than one~\cite{astrom1997computer}. 
In this paper, we express the stability specification $\phi_{stability}$ 
in terms of a check known as Jury's criterion~\cite{fadali}. 
This condition boils down to an easy algebraic formula to select the entries 
of matrix $K$, so that the closed-loop dynamics are shaped as desired. 
A detailed description of this specification can be found in the 
Appendix~\ref{sec:appendix-stability}.

@@

Furthermore, 
it is practically relevant to consider constraints on the input signal $u_{k}$
and initial states $x_0$, 
which we assume have given bounds. %%  $\overline{u}$, $\underline{u}$
%% and $\overline{x_0}$, $\underline{x_0}$, respectively.
We refer to these constraints as $\phi_{init}$:
$$\phi_{init} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} 
\wedge \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$$

This means that the control input saturates in view of physical constraints.

>>>>

We describe and evaluate two approaches for synthesising digital controllers 
for state-space physical plants. Both approaches consider a counterexample 
guided inductive synthesis (CEGIS) loop. In the first approach, 
we devise a digital controller that stabilizes the system whilst remaining safe given a 
pre-selected time horizon; then, we verify its 
potential safety by unfolding the dynamics of the system and checking for 
a completeness threshold. In the second approach, we substitute the second step by employing abstract acceleration
to unfold the dynamics forward, instead of symbolically unwinding the 
transition relation~\cite{cattaruzza2015unbounded}. 
\todo{We need to emphasise the derivation of errors for soundness. Can we claim completeness already?}
We provide experimental results 
showing that both approaches are able to efficiently synthesise stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is \textcolor{red}{$XXX$}s, i.e., most controllers can be synthesized in less
than \textcolor{red}{$XXX$} seconds.

\subsection*{Contributions}

\begin{enumerate}
\item In this paper we extend the problem of digital control synthesis from a transfer function representation into a state space representation.
This entails the verification of safety properties in addition to the already existing stability conditions which is a much harder problem.
\item To achieve this, we develop two approaches and compare them, the first relying on a non-deterministic simulation of the system and a completeness proof, whilst the second leverages on the concepts of abstraction and acceleration to increase speed whilst retaining soundness.
\item We develop a solution generalization algorithm for the first approach and an abstraction refinement for the second, both of which improve the performance of the CEGIS loop. 
\item We develop a model for different sources quantization and their effect on the state space progression, for which we find safe bounds that ensure the safety of our controllers in a hybrid continuous-digital domain.
\end{enumerate}

@@

Models \eqref{eq:ode} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t), u_k$, respectively. 
Feedback architectures can be employed to affect the properties and behaviours of the continuous process (the plant).   
%reduce the effect of parametric uncertainties and attenuate the effects of noise and disturbance. 
We are interested in the synthesis of digital feedback, 
as implemented on modern FPGAs or ICs. 
The most basic feedback architecture is the state feedback, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation}
\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation}
%
Here $K \in \mathbb{R}^{m \times n}$ is a feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital).   
%
The closed-loop model then takes the form 
\begin{align}
\label{eq:closedloopss}
x_{k+1} = ( A_d - B_d K ) x_k + B_d r_k.
\end{align}
The gain matrix $K$ can be set so that the closed-loop digitalized dynamics are shaped as desired, 
for instance according to a specific stability goal or around a specific dynamical behaviour \cite{astrom1997computer}. 
As argued later in this work, we will target more complex objectives, such as quantitative safety requirements that are not typical in the digital control literature. 
Further, we will encompass the digital nature of the controller, which manipulates quantities with finite precision.

@@

A discrete-time system as \eqref{eq:closedloopss} is asymptotically stable 
\textit{iff} all the roots of the characteristic polynomial 
(i.e., the eigenvalues of the closed-loop matrix $A_d - B_d K$) 
are inside the unity circle in the complex plane, i.e., \textit{iff} their absolute 
values are less than one~\cite{astrom1997computer}. 
In this paper, we express the stability specification $\phi_{stability}$ 
in terms of a check known as Jury's criterion~\cite{fadali}:  
this is an easy algebraic formula to select the entries of matrix $K$, 
so that the closed-loop dynamics are shaped as desired. 
A detailed description of this specification can be found in Appendix~\ref{sec:appendix-stability}.

@@

Furthermore, 
it is practically relevant to consider constraints on the input signal $u_{k}$
and bounds on the initial states $x_0$, 
which we assume have given bounds. %%  $\overline{u}$, $\underline{u}$
%% and $\overline{x_0}$, $\underline{x_0}$, respectively.
We refer to these constraints as $\phi_{init}$:
$$\phi_{init} = {\forall k.\underline{u} \leq u_{k} \leq \overline{u}} 
\wedge \bigwedge_{i=1}^{n} \underline{x_{i,0}} \leq x_{i,0} \leq \overline{x_{i,0}}.$$  
In the first case, 
this means that the control input saturates in view of physical constraints.
